---
title: "Job_Change_Analysis_of_Data_Scientists"
author: "Manisha Rana"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Objective
### A company which is active in Big Data and Data Science offers trainings to people and they want to assess the potential of hiring Data Scientists from this pool of people who successfully pass some of the courses that the company offers. Information related to demographics, education, experience and other details are collected from people when they enroll for the trainings. Based on these factors company wants to understand which people can consider joining them as Data Scientists. Target variable is the probability of a candidate to either look for a job change or not. The training set contains the target variable along with other variables and the test set does not contain the target variable.

### Importing the libraries
```{r}
library(dplyr)
library(ggplot2)
library(car)
library(GGally)
library(fastDummies)
library(caret)
library(DescTools)
library(rpart)
library(randomForest)
library(rpart)
library(e1071)
library(rattle)
library(VIM)
library(mice)
```

### Loading the train and test csv files. The dataset has been downloaded from Kaggle
```{r}
associate_train_df <- read.csv("aug_train.csv",header=TRUE,sep=",")
associate_test_df <- read.csv("aug_test.csv",header=TRUE,sep=",")
```

### Checking the structure of the data
```{r}
str(associate_train_df) 
str(associate_test_df) 
```
### the train dataset has 19158 observations and 14 variables. The test dataset has 2129 observations and 13 variables

### Checking for missing data in the dataset and converting empty cells to NA
```{r}
associate_train_df[associate_train_df==""] <- NA
colSums(is.na(associate_train_df))
associate_test_df[associate_test_df==""] <- NA
colSums(is.na(associate_test_df))
```

### INSIGHTS - Gender,Enrolled_university, education_level, major discipline, experience, company size, company type, last new job columns has missing data

### Removing enroll_id and city columns from the train and test data
```{r}
associate_sub_train_df <- subset(associate_train_df,select = -c(enrollee_id,city))
associate_sub_test_df <- subset(associate_test_df,select = -c(enrollee_id,city))
```

## EXPLORATORY DATA ANALYSIS
### Analyzing the gender distribution in dataset
```{r}
ggplot(associate_sub_train_df,aes(x=gender))+
  geom_bar(fill = "green")+
  xlab("Gender")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1)+
  ggtitle("Gender Distribution")

prop.table(table(associate_sub_train_df$gender)) 
```
### Basis gender data is distributed as 90% males,8% fenmales and 1% others

### Analysing the distribution by experience in dataset
```{r}
ggplot(associate_sub_train_df,aes(x=experience))+
  geom_bar(fill = "blue")+
  xlab("Experience")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1)+
  ggtitle("Experience Distribution")

prop.table(table(associate_sub_train_df$experience))
```
### Maximum number of employees (17%) has experience>20yrs

## Analysing Distribution by relevant experience
```{r}
ggplot(associate_sub_train_df,aes(x=relevent_experience))+
  geom_bar(fill = "blue")+
  xlab("Relevant Experience")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1)+
  ggtitle("Relevant Experience Distribution")

prop.table(table(associate_sub_train_df$relevent_experience)) 
```
### 72% of the trainees have relevant experience and 28% have no relevant experience


### Analysing Distribution by enrolled university
```{r}
ggplot(associate_sub_train_df,aes(x=enrolled_university))+
  geom_bar(fill = "blue")+
  xlab("Enrolled University")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1)+
  ggtitle("Distribution by Enrolled University")

prop.table(table(associate_sub_train_df$enrolled_university))
```
### 74% of the trainees are currently not enrolled in university, 20% are enrolled in a full time course and 6% in part time courses

### Analysing distribution by Education Level
```{r}
ggplot(associate_sub_train_df,aes(x=education_level))+
  geom_bar(fill = "blue")+
  xlab("Education")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1)+
  ggtitle("Distribution by Education Level")

prop.table(table(associate_sub_train_df$education_level))
```
## Graduates form the highest proportion (62%) of trainees who enroll for trainings followed by trainees with 23% holding Masters degree, 10% holding High School degree, 2% holding Phd degree and 1% holding primary school degree

### Analysing distribution by Major
```{r}
ggplot(associate_sub_train_df,aes(x=major_discipline))+
  geom_bar(fill = "blue")+
  xlab("Major")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1)+
  ggtitle("Distribution by Major")

prop.table(table(associate_sub_train_df$major_discipline))
```
### Maximum trainees have majors in STEM (89%), followed by Humanities (4%), Business Degree and others (2%), Arts and no major (1%)

### Analysing distribution by Company Size
```{r}
ggplot(associate_sub_train_df,aes(x=company_size))+
  geom_bar(fill = "blue")+
  xlab("Company Size")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1)+
  ggtitle("Distribution by Company size of Trainees")

prop.table(table(associate_sub_train_df$company_size))
```
### 23% of trainees work in company size (50-999), 19% in (10-500), 15% in (10000+), 11% in (10-49)

### Analysing distribution by Company Type
```{r}
ggplot(associate_sub_train_df,aes(x=company_type))+
  geom_bar(fill = "blue")+
  xlab("Company Type")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1)+
  ggtitle("Distribution by Company type of Trainees")

prop.table(table(associate_sub_train_df$company_type))
```
### 75% of the trainees are from Pvt Ltd, 7% each from Public and Funded startup, 4% each from NGO and Early Stage Startup

### Analysing distribution by Last New Job
```{r}
ggplot(associate_sub_train_df,aes(x=last_new_job))+
  geom_bar(fill = "blue")+
  xlab("Last New Job")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1)+
  ggtitle("Distribution by Last New Job of Trainees")

prop.table(table(associate_sub_train_df$last_new_job))
```
## 43% of the trainees have changed job in last 1 year followed by 18% who have changes last job >4 years back and 13% who have never chnaged a job

### Analysing distribution by Education_Level and Relevant Experience
```{r}
ggplot(associate_sub_train_df,aes(x=major_discipline,fill=relevent_experience))+
  geom_bar()+
  facet_wrap("gender")+
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))+
  xlab("Major")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1,check_overlap = TRUE)+
  ggtitle("Distribution by College Major and Relevant Experience")
```

### Analysing distribution by City Development Index and Relevant Experience
```{r}
ggplot(associate_sub_train_df,aes(x=city_development_index,fill=relevent_experience))+
  geom_bar()+
  xlab("City Development Index")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1,check_overlap = TRUE)+
  ggtitle("Distribution by City Development Index and Relevant Experience")
```
## More trainees have  relevant experience in cities with high city development index

## Analysing the data with respect to target

## Target vs Gender
```{r}
ggplot(associate_sub_train_df,aes(x=gender,fill=gender))+
  geom_bar()+
  facet_wrap("target")+
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))+
  xlab("gender")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1,check_overlap = TRUE)+
  ggtitle("Target by Gender")

prop.table(table(associate_sub_train_df$target,associate_sub_train_df$gender))
```
### 20% of males and 2% females have been looking for a job change

## Target and Experience
```{r}
ggplot(associate_sub_train_df,aes(x=experience,fill=experience))+
  geom_bar()+
  facet_wrap("target")+
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))+
  xlab("Experience")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1,check_overlap = TRUE)+
  ggtitle("Target by Experience")

prop.table(table(associate_sub_train_df$target,associate_sub_train_df$experience))
```
### More than 2% of trainees each with experience >20 yrs, 2-7yrs looking for job change

### Target and Relevant Experience
```{r}
ggplot(associate_sub_train_df,aes(x=relevent_experience,fill=education_level))+
  geom_bar()+
  facet_wrap("target")+
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))+
  xlab("Relevant Experience")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1,check_overlap = TRUE)+
  ggtitle("Target by Relevant Experience")

prop.table(table(associate_sub_train_df$target,associate_sub_train_df$education_level,associate_sub_train_df$relevent_experience))
prop.table(table(associate_sub_train_df$target,associate_sub_train_df$relevent_experience))
```
### 11% of the graduates,3% of the trainees with masters degree who has relevant degree looking for job change, 16% of the trainees with STEM who has relevant experience looking for a job

### Target and Major, Relevant Experience
```{r}
ggplot(associate_sub_train_df,aes(x=major_discipline,fill=relevent_experience))+
  geom_bar()+
  facet_wrap("target")+
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))+
  xlab("Major")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1,check_overlap = TRUE)+
  ggtitle("Target by Major")

prop.table(table(associate_sub_train_df$target,associate_sub_train_df$major_discipline,associate_sub_train_df$relevent_experience))
prop.table(table(associate_sub_train_df$target,associate_sub_train_df$major_discipline,associate_sub_train_df$education_level,associate_sub_train_df$relevent_experience))
```
### 11% of the graduates with STEM as major , having relevant experience looking for a job change

### Target and Last NEw job
```{r}
ggplot(associate_sub_train_df,aes(x=last_new_job,fill=relevent_experience))+
  geom_bar()+
  facet_wrap("target")+
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))+
  xlab("Last New Job")+
  ylab("Count")+
  geom_text(stat='count', aes(label=..count..), vjust=-1,check_overlap = TRUE)+
  ggtitle("Target by LAst New job")

prop.table(table(associate_sub_train_df$target,associate_sub_train_df$last_new_job,associate_sub_train_df$relevent_experience))
```
### 7% of the trainees who has changed new job in last 1 year looking for a job change

## DATA PREPROCESSING

### Data Cleaning - replacing "never" and ">4" in last new job with '0" and "5"
```{r}
associate_sub_test_df <-associate_sub_test_df %>% mutate(last_new_job = ifelse(last_new_job == "never", 0, last_new_job ))
associate_sub_test_df <-associate_sub_test_df %>% mutate(last_new_job = ifelse(last_new_job == ">4", 5, last_new_job ))

associate_sub_train_df <-associate_sub_train_df %>% mutate(last_new_job = ifelse(last_new_job == "never", 0, last_new_job ))
associate_sub_train_df <-associate_sub_train_df %>% mutate(last_new_job = ifelse(last_new_job == ">4", 5, last_new_job ))
```

### Data cleaning - replacing ">20 and <1" in experience to "21 and 0.5"
```{r}
associate_sub_test_df <-associate_sub_test_df %>% mutate(experience = ifelse(experience == "<1", 0.5, experience ))
associate_sub_test_df <-associate_sub_test_df %>% mutate(experience = ifelse(experience == ">20", 21, experience ))

associate_sub_train_df <-associate_sub_train_df %>% mutate(experience = ifelse(experience == "<1", 0.5, experience ))
associate_sub_train_df <-associate_sub_train_df %>% mutate(experience = ifelse(experience == ">20", 21, experience ))
```

### Data Cleaning - Replacing "10/49" in company size to "10-49"
```{r}
associate_sub_test_df <-associate_sub_test_df %>% mutate(company_size = ifelse(company_size == "10/49", "10-49", company_size ))
associate_sub_train_df <-associate_sub_train_df %>% mutate(company_size = ifelse(company_size == "10/49", "10-49", company_size ))
```

## Converting datatypes of variables in test and train dataset
```{r}
associate_sub_test_df$experience <- as.numeric(associate_sub_test_df$experience)
associate_sub_test_df$last_new_job<- as.numeric(associate_sub_test_df$last_new_job)
associate_sub_test_df$gender <- as.factor(associate_sub_test_df$gender)
associate_sub_test_df$relevent_experience <- as.factor(associate_sub_test_df$relevent_experience)
associate_sub_test_df$enrolled_university <- as.factor(associate_sub_test_df$enrolled_university)
associate_sub_test_df$major_discipline  <- as.factor(associate_sub_test_df$major_discipline )
associate_sub_test_df$company_type  <- as.factor(associate_sub_test_df$company_type )
associate_sub_test_df$company_size  <- as.factor(associate_sub_test_df$company_size )
associate_sub_test_df$education_level <- factor(associate_sub_test_df$education_level, ordered = TRUE, 
                                              levels = c("Phd", "Masters", "Graduate", "High School", "Primary School"))

associate_sub_train_df$experience <- as.numeric(associate_sub_train_df$experience)
associate_sub_train_df$last_new_job<- as.numeric(associate_sub_train_df$last_new_job)
associate_sub_train_df$gender <- as.factor(associate_sub_train_df$gender)
associate_sub_train_df$relevent_experience <- as.factor(associate_sub_train_df$relevent_experience)
associate_sub_train_df$enrolled_university <- as.factor(associate_sub_train_df$enrolled_university)
associate_sub_train_df$major_discipline  <- as.factor(associate_sub_train_df$major_discipline )
associate_sub_train_df$company_type  <- as.factor(associate_sub_train_df$company_type )
associate_sub_train_df$company_size  <- as.factor(associate_sub_train_df$company_size )
associate_sub_train_df$education_level <- factor(associate_sub_train_df$education_level, ordered = TRUE, 
                                              levels = c("Phd", "Masters", "Graduate", "High School", "Primary School"))
associate_sub_train_df$target  <- as.factor(associate_sub_train_df$target)
```

### Checking for proportion of missing values in Train and Test data
```{r}
table(is.na(associate_sub_train_df))
table(is.na(associate_sub_test_df))
table(is.na(associate_sub_train_df$last_new_job))
str(associate_sub_train_df)
str(associate_sub_test_df)

missing_data_plot_train <- aggr(associate_sub_train_df, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE,axes=TRUE,cex.lab=1.2,cex.axis=.7)
missing_data_plot_test <- aggr(associate_sub_test_df, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE,axes=TRUE,cex.lab=1.2,cex.axis=.7)
```
### 50% of the data is missing in both train and test dataset, maximum missing data is in company_type and size. With 50% missing dataset we simply cannot remove the missing data in the dataset

### Replacing the NA in Last New Job column with column mean
```{r}
associate_sub_train_df$last_new_job[is.na(associate_sub_train_df$last_new_job)] <- mean(associate_sub_train_df$last_new_job, na.rm = TRUE) 
associate_sub_test_df$last_new_job[is.na(associate_sub_test_df$last_new_job)] <- mean(associate_sub_test_df$last_new_job, na.rm = TRUE) 
```

### Using mode imputation for missing categorical values
```{r}
associate_sub_mode_traindf <- associate_sub_train_df
associate_sub_mode_testdf <- associate_sub_test_df

val <- unique(associate_sub_mode_traindf$company_type[!is.na(associate_sub_mode_traindf$company_type)])                   # Values in vec_miss
my_mode <- val[which.max(tabulate(match(associate_sub_mode_traindf$company_type, val)))] 
associate_sub_mode_traindf$company_type_imp <- associate_sub_mode_traindf$company_type                                    # Replicate vec_miss
associate_sub_mode_traindf$company_type_imp[is.na(associate_sub_mode_traindf$company_type_imp)] <- my_mode                     # Impute by mode
```

### Evaluating how much bias is introduced with mode imputation
```{r}
missingness <- c(rep("No Missings", 6), rep("Post Imputation", 6)) # Pre/post imputation
Category <- as.factor(rep(names(table(associate_sub_mode_traindf$company_type))))                   # Categories
Count <- c(as.numeric(table(table(associate_sub_mode_traindf$company_type))), as.numeric(table(associate_sub_mode_traindf$company_type_imp)))     

data_barplot <- data.frame(missingness, Category, Count)           
```
### We see that Pvt Ltd as a category was the mode and having replaced the missing value with the mode we have increased the bias in the data. So mode imputation will not work for this dataset

### Partitioning the train dataset into train and validation
```{r}
inTrain <- createDataPartition(y=associate_sub_train_df$target, p=0.6, list=FALSE)
train_df_train <- associate_sub_train_df[inTrain, ] 
train_df_validate <- associate_sub_train_df[-inTrain, ]
dim(train_df_train)
dim(train_df_validate)
```

### Checking MICE imputation for missing values. The MICE algorithm can impute mixes of continuous, binary, unordered categorical and ordered categorical data.

### Finding missing data pattern with MICE package
```{r}
md.pattern(associate_sub_train_df)
pMiss <- function(x){sum(is.na(x))/length(x)*100}
apply(train_df_train,2,pMiss)
apply(train_df_validate,2,pMiss)
str(associate_sub_train_df)
```
### 32% of data missing in company_type, 30% in company_size, 23% in gender, 14% in major_discipline, 2% each in education level and enrolled_university

### Imputing with MICE train,validate and test dataset
```{r}
train_df_train_impute <- mice(train_df_train, m=5, seed = 123)
train_df_validate_impute <- mice(train_df_validate,m=5,seed=123)

test_df_impute <- mice(associate_sub_test_df,m=5,seed=123)

Final_train_impute <- complete(train_df_train_impute)
Final_validate_impute <- complete(train_df_validate_impute)
Final_test_impute <- complete(test_df_impute)
```

### Cross validation using 10 fold
```{r}
trcontrol <- trainControl(method="cv",number =10)
metric <- "Accuracy"
```

## MODEL BUILDING

### Using 5 algorithms - LDA, Logistic Regression, Decision Trees, KNN and Random Forest to evaluate which model works best.

### 1) Linear Discriminant Analysis
```{r}
set.seed(10)
fit_lda <- train(target~., data=Final_train_impute, method="lda", metric=metric, trControl=trcontrol)
```


### 2) Logistic Regression
```{r}
set.seed(10)
fit_glm <- train(target~., data=Final_train_impute, method="glm", family="binomial",metric=metric, trControl=trcontrol)
```

### 3) Recursive Partioning and Regression Trees
```{r}
set.seed(10)
fit_rpart <- train(target~., data=Final_train_impute, method="rpart", metric=metric, trControl=trcontrol)
```

### 4) kNN
```{r}
set.seed(10)
fit_knn <- train(target~., data=Final_train_impute, method="knn", metric=metric, trControl=trcontrol)
```

### 5) Random Forest
```{r}
set.seed(10)
fit_rf <- train(target~., data=Final_train_impute, method="rf", metric=metric, trControl=trcontrol)
```

### Comparing accuracy of models
```{r}
results <- resamples(list(lda=fit_lda, GLM = fit_glm, DT=fit_rpart, knn=fit_knn))
summary(results)
```
## INSIGHT - Recursive Partioning and Regression Trees shows the best mean accuracy of 78%

```{r}
print(fit_rpart)
```

### Make predictions on the validate dataset using Recursive Partioning and Regression Trees Model
```{r}
prediction_trainee_model <- predict(fit_rpart, Final_validate_impute, na.action = na.pass)
confusionMatrix(prediction_trainee_model, Final_validate_impute$target)
```

### Predicting on testing dataset using Recursive Partioning and Regression Trees Model
```{r}
prediction_testdf <- predict(fit_rpart, Final_test_impute, type = "raw")
Final_test_impute <- Final_test_impute %>% select(everything()) %>% mutate(target = prediction_testdf)
summary(prediction_testdf)
```


### Writing the predicted output in csv file
```{r}
write.csv(Final_test_impute,"DSJobChange_finaloutput.csv")
```























